From OpenACC to OpenMP 4: \\ Toward Automatic Translation

### 2016-04-29 paper.tex
CCS categories

%
% The code below should be generated by the tool at
% http://dl.acm.org/ccs.cfm
% Please copy and paste the code instead of the example below. 
%
\begin{CCSXML}
<ccs2012>
 <concept>
  <concept_id>10010520.10010553.10010562</concept_id>
  <concept_desc>Computer systems organization~Embedded systems</concept_desc>
  <concept_significance>500</concept_significance>
 </concept>
 <concept>
  <concept_id>10010520.10010575.10010755</concept_id>
  <concept_desc>Computer systems organization~Redundancy</concept_desc>
  <concept_significance>300</concept_significance>
 </concept>
 <concept>
  <concept_id>10010520.10010553.10010554</concept_id>
  <concept_desc>Computer systems organization~Robotics</concept_desc>
  <concept_significance>100</concept_significance>
 </concept>
 <concept>
  <concept_id>10003033.10003083.10003095</concept_id>
  <concept_desc>Networks~Network reliability</concept_desc>
  <concept_significance>100</concept_significance>
 </concept>
</ccs2012>  
\end{CCSXML}

\ccsdesc[500]{Computer systems organization~Embedded systems}
\ccsdesc[300]{Computer systems organization~Redundancy}
\ccsdesc{Computer systems organization~Robotics}
\ccsdesc[100]{Networks~Network reliability}

@@

\keywords{ACM proceedings; \LaTeX; text tagging}

@@

The translation most similar to the OpenACC code is shown in
Figure~\ref{fig:nest}(c).  The iterations of the $i$-loop are divided among
thread blocks using the \texttt{distribute} directive, and the iterations of
the $j$-loop are divided among threads in the block using the \texttt{parallel
for} directive.
%
OpenACC's \texttt{gang} clause is roughly equivalent to OpenMP's
\texttt{distribute}---effectively, it distributes work among thread blocks.
Likewise, OpenACC's \texttt{vector} clause is roughly equivalent to OpenMP's
\texttt{parallel for} inside a \texttt{teams} region: it distributes work among
threads within a block.

@@

\newcommand{\translate}[1]{\ensuremath{\mathcal{T}[\![ #1 ]\!]}}
\begin{figure*}
\begin{align*}
%
\translate{\texttt{\#pragma acc data\ }\textit{data-clauses}} &=
   \texttt{\#pragma omp target data } \translate{\textit{data-clauses}} \\
%
\translate{\texttt{copyin(}\textit{vars}\texttt{)}} &=
   \texttt{map(to:}\textit{vars}\texttt{)} \\
%
\translate{\texttt{copyout(}\textit{vars}\texttt{)}} &=
   \texttt{map(from:}\textit{vars}\texttt{)} \\
%
\translate{\texttt{copy(}\textit{vars}\texttt{)}} &=
   \texttt{map(tofrom:}\textit{vars}\texttt{)} \\
%
\translate{\texttt{create(}\textit{vars}\texttt{)}} &=
   \texttt{map(alloc:}\textit{vars}\texttt{)} \\
%
\translate{\texttt{\#pragma acc parallel\ }\textit{par-clauses}} &=
   \left.\parbox[t]{4in}{%
   \texttt{\#pragma omp target teams\ }\translate{\textit{par-clauses}}}\right. \\
%
\translate{\texttt{num\_gangs(}\textit{n}\texttt{)}} &=
   \texttt{num\_teams(}\textit{n}\texttt{)} \\
%
\translate{\texttt{vector\_length(}\textit{n}\texttt{)}} &=
   \texttt{thread\_limit(}\textit{n}\texttt{)} \\
%
\translate{\texttt{private(}\textit{vars}\texttt{)}} &=
   \texttt{private(}\textit{vars}\texttt{)} \\
%
\translate{\texttt{reduction(}\textit{vars}\texttt{)}} &=
   \texttt{reduction(}\textit{reductions}\texttt{)} \\
%
\translate{\texttt{\#pragma acc loop\ }\textit{type}\texttt{\ }\textit{loop-clauses}} &=
   \begin{cases}
   \texttt{\#pragma omp distribute parallel for\ }\translate{\textit{loop-clauses}} \\
        \hspace*{2em} \textup{ if $\textit{type} = \texttt{gang vector}$} \vspace*{.25em} \\
   \texttt{\#pragma omp distribute\ }\translate{\textit{loop-clauses}} \\
        \hspace*{2em} \textup{ if $\textit{type} = \texttt{gang}$} \vspace*{.25em} \\
   \texttt{\#pragma omp parallel for\ }\translate{\textit{loop-clauses}} \\
        \hspace*{2em} \textup{ if $\textit{type} = \texttt{vector}$} \vspace*{.25em} \\
   \epsilon
        \hspace*{1.65em} \textup{ if $\textit{type} = \texttt{seq}$} \\
   \end{cases} \\
%
\translate{\texttt{independent}} &=
   \epsilon \\
%
\end{align*}
\caption{Translation rules from OpenACC to OpenMP~4.  The function
$\mathcal{T}$ maps phrases derived from the OpenACC subset grammar to
strings corresponding to one or more OpenMP directives.  Sequences
(e.g., \textit{par-clauses}) are translated elementwise, that is,
$\translate{t_1\ t_2\ \dots\ t_n} =
\translate{t_1}\ \translate{t_2}\ \dots\ \translate{t_n}$.}
The symbol $\epsilon$ denotes the empty string.
\label{fig:rules}
\end{figure*}

@@

We have described an algorithm (\S\ref{sec:algorithm}) and a prototype tool
that converts a subset of OpenACC (Figure~\ref{fig:grammar}) to OpenMP~4.  The
translation algorithm was designed to be simple and predicatble, since it will
necessarily be used in conjunction with manual restructuring and performance
tuning.  We evaluated our tool by converting the EPCC Level~1 OpenACC
benchmarks from OpenACC compiled by PGI C/C++ to OpenMP~4 compiled by Clang.
Both targeted an NVIDIA GPU (Tesla~K40).  To keep the translation process
transparent, the programmer was required to manually convert OpenACC
\emph{kernels} loops, although the tool was able to automatically translate the
remaining 67 of 70 OpenACC directives

>>>>

%
% The code below should be generated by the tool at
% http://dl.acm.org/ccs.cfm
% Please copy and paste the code instead of the example below. 
%
\begin{CCSXML}
<ccs2012>
<concept>
<concept_id>10010147.10010169.10010175</concept_id>
<concept_desc>Computing methodologies~Parallel programming languages</concept_desc>
<concept_significance>500</concept_significance>
</concept>
<concept>
<concept_id>10010147.10011777.10011014</concept_id>
<concept_desc>Computing methodologies~Concurrent programming languages</concept_desc>
<concept_significance>300</concept_significance>
</concept>
<concept>
<concept_id>10011007.10011006.10011073</concept_id>
<concept_desc>Software and its engineering~Software maintenance tools</concept_desc>
<concept_significance>500</concept_significance>
</concept>
</ccs2012>
\end{CCSXML}

\ccsdesc[500]{Computing methodologies~Parallel programming languages}
\ccsdesc[500]{Software and its engineering~Software maintenance tools}

@@

\keywords{accelerators; GPUs; OpenACC; OpenMP; translation}

@@

The translation most similar to the OpenACC code is shown in
Figure~\ref{fig:nest}(c).  The iterations of the $i$-loop are divided among
thread blocks using the \texttt{distribute} directive, and the iterations of
the $j$-loop are divided among threads in the block using the \texttt{parallel
for} directive.
%
OpenACC's \texttt{gang} clause is roughly equivalent to OpenMP's
\texttt{distribute}: effectively, it distributes work among thread blocks.
Likewise, OpenACC's \texttt{vector} clause is roughly equivalent to OpenMP's
\texttt{parallel for} inside a \texttt{teams} region: it distributes work among
threads within a block.

@@

\newcommand{\translate}[1]{\ensuremath{\mathcal{T}[\![ #1 ]\!]}}
\begin{figure*}
\begin{align*}
%
\translate{\texttt{\#pragma acc data\ }\textit{data-clauses}} &=
   \texttt{\#pragma omp target data } \translate{\textit{data-clauses}} \\
%
\translate{\texttt{copyin(}\textit{vars}\texttt{)}} &=
   \texttt{map(to:}\textit{vars}\texttt{)} \\
%
\translate{\texttt{copyout(}\textit{vars}\texttt{)}} &=
   \texttt{map(from:}\textit{vars}\texttt{)} \\
%
\translate{\texttt{copy(}\textit{vars}\texttt{)}} &=
   \texttt{map(tofrom:}\textit{vars}\texttt{)} \\
%
\translate{\texttt{create(}\textit{vars}\texttt{)}} &=
   \texttt{map(alloc:}\textit{vars}\texttt{)} \\
%
\translate{\texttt{\#pragma acc parallel\ }\textit{par-clauses}} &=
   \left.\parbox[t]{4in}{%
   \texttt{\#pragma omp target teams\ }\translate{\textit{par-clauses}}}\right. \\
%
\translate{\texttt{num\_gangs(}\textit{n}\texttt{)}} &=
   \texttt{num\_teams(}\textit{n}\texttt{)} \\
%
\translate{\texttt{vector\_length(}\textit{n}\texttt{)}} &=
   \texttt{thread\_limit(}\textit{n}\texttt{)} \\
%
\translate{\texttt{private(}\textit{vars}\texttt{)}} &=
   \texttt{private(}\textit{vars}\texttt{)} \\
%
\translate{\texttt{reduction(}\textit{vars}\texttt{)}} &=
   \texttt{reduction(}\textit{reductions}\texttt{)} \\
%
\translate{\texttt{independent}} &=
   \epsilon \\
%
\translate{\texttt{\#pragma acc loop\ }\textit{type}\texttt{\ }\textit{loop-clauses}} &=
   \begin{cases}
   \texttt{\#pragma omp distribute parallel for\ }\translate{\textit{loop-clauses}} \\
        \hspace*{2em} \textup{ if $\textit{type} = \texttt{gang vector}$} \vspace*{.25em} \\
   \texttt{\#pragma omp distribute\ }\translate{\textit{loop-clauses}} \\
        \hspace*{2em} \textup{ if $\textit{type} = \texttt{gang}$} \vspace*{.25em} \\
   \texttt{\#pragma omp parallel for\ }\translate{\textit{loop-clauses}} \\
        \hspace*{2em} \textup{ if $\textit{type} = \texttt{vector}$} \vspace*{.25em} \\
   \epsilon
        \hspace*{1.65em} \textup{ if $\textit{type} = \texttt{seq}$} \\
   \end{cases} \\
%
\end{align*}
\caption{Translation rules from OpenACC to OpenMP~4.  The function
$\mathcal{T}$ maps phrases derived from the OpenACC subset grammar to
strings corresponding to one or more OpenMP directives.  Sequences
(e.g., \textit{par-clauses}) are translated elementwise, that is,
$\translate{t_1\ t_2\ \dots\ t_n} =
\translate{t_1}\ \translate{t_2}\ \dots\ \translate{t_n}$.}
The symbol $\epsilon$ denotes the empty string.
\label{fig:rules}
\end{figure*}

@@

We have described an algorithm (\S\ref{sec:algorithm}) and a prototype tool
that converts a subset of OpenACC (Figure~\ref{fig:grammar}) to OpenMP~4.  The
translation algorithm was designed to be simple and predicatble, since it will
necessarily be used in conjunction with manual restructuring and performance
tuning.  We evaluated our tool by converting the EPCC Level~1 OpenACC
benchmarks from OpenACC compiled by PGI C/C++ to OpenMP~4 compiled by Clang.
Both targeted an NVIDIA GPU (Tesla~K40).  To keep the translation process
transparent, the programmer was required to manually convert OpenACC
\emph{kernels} loops, although the tool was able to automatically translate the
remaining 67 of 70 directives.
